---
output: html_notebook
---

### Load the packages

```{r setup}
library(tidyverse)
library(sf)
library(spdep)
library(purrr)
library(magrittr)
library(forcats)


```

### Load the data

```{r}
nc <- st_read(system.file("shape/nc.shp", package="sf")) 

nc_shp <- as(nc, 'Spatial')
```



### Spatial clustering test
Steps:

  1. Create a new variable (`NW_PCT`) to test for spatial clustering using a local indicator of spatial association (this example will use Local Moran I)
  2. Find neighbors using rook's case contiguity
  3. Calculate the local moran I
    * save the statistic itself (`LOCAL_M_I`) and its p-value (`LOCAL_M_P`) 
  4. Categorize the clusters into four "quadrants" according to two variables:
    * their relationship to the mean `NW_PCT` value (`HIGH` if greater than, `LOW` is less than)
    * their relationship to the mean `LOCAL_M_I` value (`HIGH` if greater than, `LOW` is less than)
  5. Categorize the clusters with local Moran p-value less than 0.1 as `INSIGNIFICANT`

```{r}
nc %<>% mutate(NW_PCT = NWBIR74/BIR74)  

nb_list <- nb2listw(poly2nb(nc_shp,queen = FALSE))

extract_localmoran <- function(x,y,stat_name){ 
        localmoran(x,y) %>% as_data_frame %>% extract2(stat_name) %>% unclass %>% as.numeric()
        }

nc_local_moran <- 
        nc %>%  
        mutate(LOCAL_M_I = extract_localmoran(NW_PCT, nb_list, 'Ii'),
               LOCAL_M_P = extract_localmoran(NW_PCT, nb_list, 'Pr(z > 0)'),
               QUADRANTS = case_when(NW_PCT >= mean(NW_PCT) & LOCAL_M_I >= mean(NW_PCT) ~ 'HIGH/HIGH',
                                    NW_PCT >= mean(NW_PCT) & LOCAL_M_I < mean(NW_PCT) ~ 'HIGH/LOW',
                                    NW_PCT < mean(NW_PCT) & LOCAL_M_I >= mean(NW_PCT) ~ 'LOW/HIGH',
                                    TRUE ~ 'LOW/LOW'),
               QUADRANTS = if_else(LOCAL_M_P > 0.1, 'INSIGNIFICANT', QUADRANTS),
               QUADRANTS = factor(QUADRANTS, 
                                 ordered = TRUE,
                                 levels = c(
                                         'HIGH/HIGH',
                                         'HIGH/LOW',
                                         'LOW/HIGH',
                                         'LOW/LOW'
                                 )),
               QUADRANTS = fct_rev(QUADRANTS)) 



```

### Plot the results

```{r}

nc_local_moran %>% 
        select(NW_PCT, LOCAL_M_I, QUADRANTS) %>% 
        plot

```








