---
title: "ACS Objects in Dataframes"
author: Tiernan Martin
date: "2016-12-12"
output: html_notebook
---

### Introduction
This is an exploration of how to work with `acs` objects, which are `S4` and come from the `acs` [package](https://cran.r-project.org/web/packages/acs/acs.pdf), inside a dataframe. This exploration is inspired by the approach suggested by Jenny Bryan in her [Plotcon 2016 presentation](https://www.youtube.com/watch?v=4MfUCX_KpdE), which highlights some of the benefits of keeping related objects in a dataframe.


```{r setup}
library(acs)
library(tidyverse)

```

### Test Data: US States
From the US Census's API, download a few tables from different years or of differeing span lengths. _Note: you'll need an API key in order to connect to server. Keys are free and available [here](http://api.census.gov/data/key_signup.html)._

```{r download-data,echo=TRUE,cache=TRUE}
race_2015 <- acs::acs.fetch(endyear = 2015,span = 1,geography = geo.make(state = '*'),table.number = "B03002",col.names = 'pretty')
glimpse(race_2015)
```

### New Variable: `POC_COUNT` (People of Color)
The first method shows the way that the `acs` objects are intended to be combined and manipulated, as described in the package [supplementary documentation](http://eglenn.scripts.mit.edu/citystate/wp-content/uploads/2013/06/wpid-working_with_acs_R3.pdf).

```{r poc, echo=TRUE}

# Find the column index for all people (total)

total <- acs.colnames(race_2015)[[which(str_detect(acs.colnames(race_2015),'Total:$'))]]

# Find the column index for non-hispanic whites

white_nonhisp <- acs.colnames(race_poc_2015)[[which(str_detect(acs.colnames(race_2015),'Not Hispanic or Latino: White alone$'))]]


# Create the new column: 'POC_COUNT'
poc_2015 <- race_2015[,total] - race_2015[,white_nonhisp]
acs.colnames(poc_2015) <- 'POC_COUNT'

# Combine the new acs object with the original and print a summary

race_poc_2015 <- cbind.acs(race_2015,poc_2015)

summary(race_poc_2015[,c(total,white_nonhisp,'POC_COUNT')])

```


The second method shows how the `acs` object can be split into observations and rows, and then stored in a dataframe so that it can be manipulated with and combined.

```{r poc-df, echo=TRUE}

race_poc_2015_df <- 
        race_poc_2015 %>% 
        apply(2,function(x){apply(x,1,list) %>% as_tibble}) %>% 
        as_tibble()
colnames(race_poc_2015_df) <- acs.colnames(race_poc_2015)


race_poc_2015_df  <-  
        race_poc_2015_df %>% 
        select(TOTAL_COUNT = matches(total),
               WHITE_COUNT = matches(white_nonhisp),
               POC_COUNT) %>% 
        mutate(POC_COUNT_DF = map2(TOTAL_COUNT,WHITE_COUNT, ~ .x - .y))
        
race_poc_2015_df %>% head() %>% print(n = 5)
```

When checked, it is clear that `POC_COUNT` and `POC_COUNT_DF` are identical:
```{r}
# Comparison

compare_acs <- 
        race_poc_2015_df %>% 
        select(matches('POC')) %>%      # select the desired columns
        map(.f = function(x)unlist(x) %>% Reduce('rbind.acs',x = .)) %>%   # transform each column into an acs object      
        unlist() %>% 
        Reduce('cbind.acs',x = .)       # combine the acs objects into one acs object
print(compare_acs[1:10,])

comp_df <- data.frame(estimate(compare_acs[,1]),estimate(compare_acs[,2]))

all(comp_df[[1]] == comp_df[[2]])
```

